\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2019-06-20 draft.tex
with fullpage and times on
Switched off the full page mode
some shorten (unfinished)

%\usepackage{latexsym}
\usepackage{setspace}
\usepackage{cancel}
\usepackage{graphicx}
\usepackage{appendix}
\usepackage{amssymb}
\usepackage{times}
\usepackage{fullpage}

@@

\subsection*{Related Works}
Reducer commutativity problem can be reduced to a program equivalence problem. One creates another program $R'$ that first non-deterministically swaps two consecutive input values and then executes the code of $R$. If $R'$ and $R$ are equivalent, using the fact that all permutations of a list can be obtained by swapping consecutive list elements finitely many times\footnote{Here is an example to produce [3;2;5;1;4] from [1;2;3;4;5] by swapping consecutive elements: [\textbf{1};\textbf{2};3;4;5] $\rightarrow$ [2;\textbf{1};\textbf{3};4;5] $\rightarrow$ [2;3;1;\textbf{4};\textbf{5}]$\rightarrow$ [2;3;\textbf{1};\textbf{5};4]$\rightarrow$ [\textbf{2};\textbf{3};5;1;4]$\rightarrow$ [3;2;5;1;4].}, $R$ can be proved to be commutative. There is a series of research results on program equivalence checking (or closely related topics such as regression verification and translation validation)~\cite{Pnueli:1998:TV,symdiff,fedyukovich2015automated,sharma2013data,godlin2009regression,fedyukovich2016property,felsing2014automating,KieferKlebanovUlbrich2017,lahiri2013differential,grossman2017verifying,barthe2011relational,KlebanovRuemmerUlbrich2017}. At a high level, checking equivalence of two programs $P$ and $P'$ can be reduced to a sequential verification problem by executing $P'$ after $P$ and at the end checking if they always have the same outputs. It can be made more efficient by finding the right~\emph{synchronization points} and combining the code of $P$ and $P'$ in an interleaved manner. A lot of research effort in this direction is in finding good synchronization points. For the case of commutativity analysis, one can simply use the beginning of the loop as the synchronization point. However, in our experience, if we naively reduce the commutativity problem to equivalence problem and check it in a precise manner, many examples cannot be solved. In this paper, we propose a good over-approximation (c.f. Section~\ref{sec:prover}) of the reducer's behavior. The approximation allows a much more efficient yet precise enough commutativity analysis.

One main motivation of commutativity analysis is to reduce the verification problem of a concurrent program to a sequential one. There are several pieces of research sharing the similar motivation, e.g., the concept of \emph{robustness} of event-driven asynchronous programs~\cite{ahmed2017:robustness}, the \emph{robustness} of program running under relaxed memory model~\cite{ahmed2013:robustness,AbdullaACLR13,AbdullaACLR12}, and \emph{synchronizability} of communicating finite state machines~\cite{FinkelL17,basu2012synchronizability,basu2011choreography}. However, their computation models are quite different from that of reducers and hence the result cannot be directly applied. Besides verification, another interesting research direction is the synthesis of MapReduce programs~\cite{SmithA16}. Commutativity analysis is used as a component under the synthesis framework.

@@

\begin{equation*}
\begin{array}{rclcl}
\multicolumn{3}{l}{v\in \Var\mbox{, and }c \in \Z}&&\textmd{Variable and Constant}\\
e \in \Exp    &\ \ \  {\triangleq}\ \ \   & v \mid c \mid \cur \mid e + e \mid e - e \mid e \times e \mid *\mid\ldots & &\textmd{Integer Expression}\\
g\in \Grd  & {\triangleq}  & e < e \mid not\ g \mid g \wedge g  \mid \ldots& & \textmd{Guard}\\
s\in \Cmd  & {\triangleq}  & v:= e \mid out(v) \mid \ite{g}{s}{s} \mid s;s && \textmd{Command}\\
p\in \Prg  & {\triangleq}  & s;\rloop\{s\};s&& \textmd{Reducer Program}\\
\end{array}
\end{equation*}

Here any $v\in \Var$ is a variable and any $c \in \Z$ is a constant integer value. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.
A \emph{guard} in $\Grd$ can be a comparison between two expressions or boolean combinations of several guards. A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of a sequence of commands, or $out(v)$ that outputs the value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$, and $s_2$ is assumed to have at most one $\cur$ expression in each \emph{branch}\footnote{We define a \emph{branch} of a command $s$ as a sequence obtained by replacing every branch statement $\ite{g}{s_1}{s_2}$ in $s$ with either $(g); s_1$ or $(not\ g) ;s_2$. For example, the loop body of Figure~\ref{fig:reducer_example}(a) has two branches $t:=\cur; (t>0); m:=t$ and $t:=\cur;(not\ t>0)$ }. According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. A few examples of reducer programs in this model can be found in Figure~\ref{fig:reducer_example}.

Some reducer uses two loops to compute the output. For example, for calculating standard deviation, one first computes the average of inputs and then uses it to compute the standard deviation. In that case, we can still verify the reducer by first partitioning it into two reducers and then verifying them separately. One reducer computes the average (Figure~\ref{fig:reducer_example}(b)) and the other reducer uses a non-deterministic value to over-approximate the computed average ($a := *$) (Figure~\ref{fig:reducer_example}(c)). The reducer in Figure~\ref{fig:reducer_example}(c) ``over-approximates''  all possible outputs of the standard deviation computation; it computes an output even if $a$ is not the average of inputs. However, the reducer is still commutative and its commutativity verification can still be done.

\begin{figure}
	\begin{minipage}{0.32\textwidth}
		\begin{algorithm}[H]
			$m := \cur$; \;
			\Loop{}{
				$t:=\cur$;\;
				\uIf{ $t > m$}{
					$m := t$
				}
			};
			$out(m)$\;
		\end{algorithm}
		\caption*{(a) max}
	\end{minipage}
	\hspace{-0.5cm}
	\begin{minipage}{0.3\textwidth}
		\begin{algorithm}[H]
			$s := 0;c:=0$; \;
			\Loop{}{
				$s := s+\cur$;\;
				$c := c+1$
			};
			$o := s/c$;\;
			$out(o)$\;
		\end{algorithm}
		\caption*{(b) average}
	\end{minipage}
	\begin{minipage}{0.4\textwidth}
		\begin{algorithm}[H]
			$a:=*;s := 0; c:=0$;\;
			\Loop{}{
				$s := s+(\cur-a)^2$; \;
				$c := c+1$
			};
			$o := s/c$;\;
			$out(o)$
		\end{algorithm}
		\caption*{(c) standard deviation}
	\end{minipage}
	\caption{Example of reducer programs}
	\label{fig:reducer_example}
\end{figure}

In the next section, we give a brief overview of J-ReCoVer,  our tool for reducer commutativity verification.

@@

\begin{figure}[h]
\centering
	\scalebox{0.9}{
		\begin{tikzpicture}

		%J-ReCoVer
		\node[draw=red!30, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=9cm,fill=red!10,minimum height=8cm] (jrecover) at (5.5,0.8) {};
		\node at (5.5,0.4) {\textbf{J-ReCoVer}};

		%JAVA reducer
		\node[draw=yellow!80, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=1cm,fill=yellow!20,minimum height=4cm] (reducer) at (0,-1.6) {};
		\node[rotate=270, align = center] at (0,-3.6) {Java Reducer};

@@

		%JProver
		\node[draw=blue!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=3cm,fill=blue!20,minimum height=5.5cm] (prover) at (7,0) {};
		\node[] at (7,-0.4) {Prover};

		%JProver->Technologies
		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1.2cm] (z3) at (7,-1) {};

		\node at (7,-1.6) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Z3 SMT \\
			Solver~\cite{z3}
			\end{minipage}
		};

		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1.2cm] (z3) at (7,-2.5) {};

		\node at (7,-3.1) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Symbolic \\
			Execution
			\end{minipage}
		};

		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1.2cm] (z3) at (7,-4.) {};

		\node at (7,-4.6) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Data Flow \\
			Analysis
			\end{minipage}
		};

		%JBugFinder
		\node[draw=blue!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=3cm,fill=blue!20,minimum height=1cm] (learner) at (7,-6) {};
		\node[] at (7,-6.5) {BugFinder};

@@

		\node at (12,-1) {
			Commutative
		};

		\node at (12,-3.5) {
			Unknown
		};

		\node at (12,-6.5) {
			\begin{minipage}[t]{0.25\textwidth}
			\centering
			Non-Commutative	 \\
			+\\
			Counterexample
			\end{minipage}

@@

		\draw [->] (8.5,-6.8) -- (10.5,-6.8) ;
		\draw [->] (8.5,-1) -- (10.5,-1) ;
		\draw [-] (8.5,-6.2) -- (9.5,-6.2) --(9.5,-3.5);
		\draw [->] (8.5,-3.5) -- (10.5,-3.5) ;

		\draw [->] (4.5,-3) -- (5.5,-3) ;
		\draw [->] (4.5,-6.5) -- (5.5,-6.5) ;

		\draw [->] (0.5,-4) -- (1.5,-4) ;

		\end{tikzpicture}
	}
	\caption{Overview of the J-ReCoVer Tool. }\label{fig:overview-jrecover}
\end{figure}

@@

\begin{figure}[htb]
	\centering
	\begin{minipage}{0.8\textwidth}

	\begin{algorithm}[H]
		$s'_0;s:=*;$\;
		\Loop{}{
			\lIf{$s=1$}{ $x_{i_1} := \cur; s'_1; s:=2$}
			\lElseIf{ $s =2$}{ $x_{i_2} := \cur; s'_2; s:=3$}
			\lElseIf{ $s = 3$}{$\ldots$}
			\lElseIf{ $s = m$}{$x_{i_m} := \cur; s'_m; s:=m+1$}
			\lElse{ $s_2$}
		};$s_3$\;
	\end{algorithm}
	\end{minipage}
	\caption{The second transformation task}
	\label{fig:general_program_transformation}
\end{figure}

>>>>

%\usepackage{latexsym}
\usepackage{setspace}
\usepackage{cancel}
\usepackage{graphicx}
\usepackage{appendix}
\usepackage{amssymb}
\usepackage{times}
% \usepackage{fullpage}

@@

\subsection*{Related Works}
Reducer commutativity problem can be reduced to a program equivalence problem. One creates another program $R'$ that first non-deterministically swaps two consecutive input values and then executes the code of $R$. If $R'$ and $R$ are equivalent, using the fact that all permutations of a list can be obtained by swapping consecutive list elements finitely many times\footnote{Here is an example to produce [3;2;5;1;4] from [1;2;3;4;5] by swapping consecutive elements: [\textbf{1};\textbf{2};3;4;5] $\rightarrow$ [2;\textbf{1};\textbf{3};4;5] $\rightarrow$ [2;3;1;\textbf{4};\textbf{5}]$\rightarrow$ [2;3;\textbf{1};\textbf{5};4]$\rightarrow$ [\textbf{2};\textbf{3};5;1;4]$\rightarrow$ [3;2;5;1;4].}, $R$ can be proved to be commutative. There is a series of research results on program equivalence checking (or closely related topics such as regression verification and translation validation), to name a few~\cite{Pnueli:1998:TV,fedyukovich2015automated,barthe2011relational,KlebanovRuemmerUlbrich2017}. At a high level, checking equivalence of two programs $P$ and $P'$ can be reduced to a sequential verification problem by executing $P'$ after $P$ and at the end checking if they always have the same outputs. It can be made more efficient by finding the right~\emph{synchronization points} and combining the code of $P$ and $P'$ in an interleaved manner. A lot of research effort in this direction is in finding good synchronization points. For the case of commutativity analysis, one can simply use the beginning of the loop as the synchronization point. However, in our experience, if we naively reduce the commutativity problem to equivalence problem and check it in a precise manner, many examples cannot be solved. In this paper, we propose a good over-approximation (c.f. Section~\ref{sec:prover}) of the reducer's behavior. The approximation allows a much more efficient yet precise enough commutativity analysis.

One main motivation of commutativity analysis is to reduce the verification problem of a concurrent program to a sequential one. There are several pieces of research sharing the similar motivation, e.g., the concept of \emph{robustness} of event-driven asynchronous programs~\cite{ahmed2017:robustness} and program running under relaxed memory model~\cite{ahmed2013:robustness,AbdullaACLR13,AbdullaACLR12}. However, their computation models are quite different from that of reducers and hence the result cannot be directly applied. Besides verification, another interesting research direction is the synthesis of MapReduce programs~\cite{SmithA16}. Commutativity analysis is used as a component under the synthesis framework.

@@

\hide{
\begin{equation*}
\begin{array}{rclcl}
\multicolumn{3}{l}{v\in \Var\mbox{, and }c \in \Z}&&\textmd{Variable and Constant}\\
e \in \Exp    &\ \ \  {\triangleq}\ \ \   & v \mid c \mid \cur \mid e + e \mid e - e \mid e \times e \mid *\mid\ldots & &\textmd{Integer Expression}\\
g\in \Grd  & {\triangleq}  & e < e \mid not\ g \mid g \wedge g  \mid \ldots& & \textmd{Guard}\\
s\in \Cmd  & {\triangleq}  & v:= e \mid out(v) \mid \ite{g}{s}{s} \mid s;s && \textmd{Command}\\
p\in \Prg  & {\triangleq}  & s;\rloop\{s\};s&& \textmd{Reducer Program}\\
\end{array}
\end{equation*}}
\begin{wrapfigure}{r}{3cm}
	\hide{
		\begin{minipage}{0.32\textwidth}
			\begin{algorithm}[H]
				$m := \cur$; \;
				\Loop{}{
					$t:=\cur$;\;
					\uIf{ $t > m$}{
						$m := t$
					}
				};
				$out(m)$\;
			\end{algorithm}
			\caption*{(a) max}
		\end{minipage}
		\hspace{-0.5cm}
	}
	\begin{minipage}{0.3\textwidth}
		\begin{algorithm}[H]
			$s := 0;c:=0$; \;
			\Loop{}{
				$s := s+\cur$;\;
				$c := c+1$
			};
			$o := s/c$;\;
			$out(o)$\;
		\end{algorithm}
	\end{minipage}
	\hide{
		\begin{minipage}{0.4\textwidth}
			\begin{algorithm}[H]
				$a:=*;s := 0; c:=0$;\;
				\Loop{}{
					$s := s+(\cur-a)^2$; \;
					$c := c+1$
				};
				$o := s/c$;\;
				$out(o)$
			\end{algorithm}
			\caption*{(c) standard deviation}
		\end{minipage}
	}
	\caption{Average}
	\label{fig:reducer_example}
\end{wrapfigure}
Let $\Var$ be the set of variables and $\Z$ the set of integers. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.

A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of a sequence of commands, or $out(v)$ that outputs the value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$, and $s_2$ is assumed to have at most one $\cur$ expression in each \emph{branch}\footnote{We define a \emph{branch} of a command $s$ as a sequence obtained by replacing every branch statement $\ite{g}{s_1}{s_2}$ in $s$ with either $(g); s_1$ or $(not\ g) ;s_2$. For example, the loop body of Figure~\ref{fig:reducer_example}(a) has two branches $t:=\cur; (t>0); m:=t$ and $t:=\cur;(not\ t>0)$ }. According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. One example can be found in Figure~\ref{fig:reducer_example}.

\hide{
Some reducer uses two loops to compute the output. For example, for calculating standard deviation, one first computes the average of inputs and then uses it to compute the standard deviation. In that case, we can still verify the reducer by first partitioning it into two reducers and then verifying them separately. One reducer computes the average (Figure~\ref{fig:reducer_example}(b)) and the other reducer uses a non-deterministic value to over-approximate the computed average ($a := *$) (Figure~\ref{fig:reducer_example}(c)). The reducer in Figure~\ref{fig:reducer_example}(c) ``over-approximates''  all possible outputs of the standard deviation computation; it computes an output even if $a$ is not the average of inputs. However, the reducer is still commutative and its commutativity verification can still be done.

In the next section, we give a brief overview of J-ReCoVer,  our tool for reducer commutativity verification.}

@@

\begin{wrapfigure}{r}{7cm}
	\vspace{-0.5cm}
\centering
	\scalebox{0.7}{
		\begin{tikzpicture}

		%J-ReCoVer
		\node[draw=red!30, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=8cm,fill=red!10,minimum height=8cm] (jrecover) at (5,0.8) {};
		\node at (5,0.4) {\textbf{J-ReCoVer}};

		%JAVA reducer
		\node[draw=yellow!80, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=0.6cm,fill=yellow!20,minimum height=4cm] (reducer) at (0.5,-1.6) {};
		\node[rotate=270, align = center] at (0.5,-3.6) {Java Reducer};

@@

		%JProver
		\node[draw=blue!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=3cm,fill=blue!20,minimum height=5.5cm] (prover) at (6.5,0) {};
		\node[] at (6.5,-0.4) {Prover};

		%JProver->Technologies
		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1.2cm] (z3) at (6.5,-1) {};

		\node at (6.5,-1.6) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Z3 SMT \\
			Solver~\cite{z3}
			\end{minipage}
		};

		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1.2cm] (z3) at (6.5,-2.5) {};

		\node at (6.5,-3.1) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Symbolic \\
			Execution
			\end{minipage}
		};

		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1.2cm] (z3) at (6.5,-4.) {};

		\node at (6.5,-4.6) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Data Flow \\
			Analysis
			\end{minipage}
		};

		%JBugFinder
		\node[draw=blue!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=3cm,fill=blue!20,minimum height=0.5cm] (learner) at (6.5,-6.2) {};
		\node[] at (6.5,-6.5) {BugFinder};

@@

		\node[rotate=270, align = center] at (9.5,-1) {
			Commutative
		};

		\node[rotate=270, align = center] at (9.5,-3.5) {
			Unknown
		};

		\node[rotate=270, align = center] at (9.6,-6) {
			\begin{minipage}[t]{0.25\textwidth}
			\centering
			Non-Commutative	 \\
			+ Counterexample
			\end{minipage}

@@

		\draw [->] (8,-6.6) -- (9.2,-6.6) ;
		\draw [->] (8,-1) -- (9.2,-1) ;
		\draw [-] (8,-6.3) -- (8.5,-6.3) --(8.5,-3.5);
		\draw [->] (8,-3.5) -- (9.2,-3.5) ;

		\draw [->] (4.5,-3) -- (5,-3) ;
		\draw [->] (4.5,-6.5) -- (5,-6.5) ;

		\draw [->] (0.8,-4) -- (1.5,-4) ;

		\end{tikzpicture}
	}
		\vspace{-1cm}
	\caption{Overview of the J-ReCoVer Tool. }\label{fig:overview-jrecover}
		\vspace{-0.8cm}

\end{wrapfigure}

@@

\begin{wrapfigure}{r}{9cm}
	\centering
	\begin{minipage}{0.8\textwidth}

	\begin{algorithm}[H]
		$s'_0;s:=*;$\;
		\Loop{}{
			\lIf{$s=1$}{ $x_{i_1} := \cur; s'_1; s:=2$}
			\lElseIf{ $s =2$}{ $x_{i_2} := \cur; s'_2; s:=3$}
			\lElseIf{ $s = 3$}{$\ldots$}
			\lElseIf{ $s = m$}{$x_{i_m} := \cur; s'_m; s:=m+1$}
			\lElse{ $s_2$}
		};$s_3$\;
	\end{algorithm}
	\end{minipage}
	\caption{The second transformation task}
	\label{fig:general_program_transformation}
\end{wrapfigure}
